---
title: "Template for Project 3 Data Analysis Report (replace it with your own title)"
author: "Liuyi Bao, ID:918968849;Jun Wei,ID: 918978635;Xiaoyi Xu, ID:918962492; Yunhao Yang,ID:918347075; Team 1"
date: "(Date on which the report is written)"
output:
  html_document:
    df_print: paged
    number_sections: yes
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

```{r, message=FALSE, warning=FALSE,include=FALSE}
library(RJSONIO)
library(tidyverse)
library(lubridate)
library(plotly)
library(echarts4r)
library(DT)
library(sparkline)
library(XML)
library(RCurl)
library(gplots)
library(tidyverse)
library(dplyr)
library(RJSONIO)
library(tidyverse)
library(lubridate)
library(plotly)
library(echarts4r)
library(DT)
library(qwraps2)
library(ggplot2)
library(lmerTest)
library(maps)

```

```{r,echo=FALSE,message=F}

covid <- read_csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")
```


```{r,echo=FALSE,message=FALSE}
covidmonth<-covid
covidmonth$month<-format(as.Date(covid$Date_reported, "%Y-%m-%d"), "%Y-%m")
covidmonth$Country[covidmonth$Country =="United States of America"]<-"United States"
covidmonth$Country[covidmonth$Country =="Russian Federation"]<-"Russia"
covidmonth$Country[covidmonth$Country =="Iran (Islamic Republic of)"]<-"Iran"
covidmonth$Country[covidmonth$Country =="Congo"]<-"Dem. Rep. Congo"
covidmonth$Country[covidmonth$Country =="United Republic of Tanzania"]<-"Tanzania"
covidmonth$Country[covidmonth$Country =="South Sudan"]<-"S. Sudan"
covidmonth$Country[covidmonth$Country =="Central African Republic"]<-"Central African Rep."
covidmonth$Country[covidmonth$Country =="Venezuela (Bolivarian Republic of)"]<-"Venezuela"
covidmonth$Country[covidmonth$Country =="Bolivia (Plurinational State of)"]<-"Bolivia"
covidmonth$Country[covidmonth$Country =="Viet Nam"]<-"Vietnam"
```

```{r,echo=FALSE,message=FALSE}

covidmonth = covidmonth %>% filter(month != "2021-03")
covidmonth1<-covidmonth %>% group_by(Country,month) %>%
mutate(
  New_cases=sum(New_cases),   Cumulative_cases=sum(Cumulative_cases), 
  New_deaths=sum(New_deaths),
  Cumulative_deaths=sum(Cumulative_deaths))

data<-covidmonth1[, !(colnames(covidmonth1) %in% c("Date_reported"))]%>%unique()%>%select(month, everything())


#data%>% filter(Country_code=="US")
c_data = covidmonth %>% filter(Date_reported == "2021-02-28") 
c_data = c_data[,c(3,6,8)]%>%distinct() %>% mutate(death_rate = Cumulative_deaths/Cumulative_cases)
```


```{r,echo=FALSE}
data_region<-covidmonth %>%  filter(WHO_region != "Other") %>% group_by(WHO_region,month) %>%
mutate(RNew_cases=sum(New_cases),   RCumulative_cases=sum(Cumulative_cases), 
  RNew_deaths=sum(New_deaths),
  RCumulative_deaths=sum(Cumulative_deaths))%>%
  select(month,WHO_region,RNew_cases,RNew_deaths,RCumulative_cases,RCumulative_deaths)%>% distinct() %>% filter(month != "2021-03") %>% mutate(deathrate = RCumulative_deaths/RCumulative_cases)

```





***

# Abstract 


# Introduction
## Motivation of analysis
Due to COVID-19, the whole world is suffering from a burdensome pandemic.Not only millions of people had been infected, but also more than 2 millions people lost their lives in this pandemic.Up to now, many countries in the world still face the huge pressure caused by the epidemic, not only in the aspect of public health but also in the aspects of economic and politics. To help the public recognize the current situation of epidemic, we will try to depict this pandemic through data and try to find some causal relationship about deaths and the number of people infected. We hope that our project can be the reference for the normal people to understand the epidemic more objectively and help the government and policymakers to face the epidemic easier.   

## Question of interest

In this project, we will focus on the data about number of the confirmed cases and number of deaths including accumulating number of cases and deaths and new cases and deaths every day. To help the public understand the current situation of the epidemic and find some causal relationship about the data, we will try to answer several questions we are interested in in this project.   
  
1. How the number of new cases and new deaths change with the time in each region and the whole world? How can we  evaluate this trend?   
  
2. Whether there is any differences in new cases and new deaths each month between different regions?   
  
3. How can we describe the relationship between new cases and new deaths? 
  
  


## Potential impact of result 
The result of our project may be a good reference for the public to recognize the current situation of the epidemic and help them know how the epidemic developed in the past time. Moreover, our project also can help the government make policy to control the epidemic and improve the sitiuation of public health.
 
# Background 



# Descriptive analysis 
As we explained before, we will not analysis the data set day by day but we will concern monthly data instead. First, we can check the summary of monthly new cases, new deaths, cumulative cases and cumulative deaths.
```{r,results='asis',message=FALSE, warning=FALSE,echo=F}
# summary table
options(qwraps2_markup = 'markdown')
Summarystat <-qsummary(covidmonth[,c('New_cases','Cumulative_cases','New_deaths','Cumulative_deaths')],n_perc_args = list(digits = 1))
summary_table(dplyr::group_by(covidmonth, month),Summarystat)
```
In summary table, some minimum of new cases and new deaths are negative, probably because the cases and deaths were wrongly attributed to the Covid-19 before and the data was modified when mistake was found. The median and mean of new cases were increasing from 2020/01 to 2021/01, but they were both decreasing from 2021/01 to 2021/02. The standard deviation of new cases peaked at 14731.94 in 2021/01. The mean of new death showed a general tendancy to increase from 2020/01 to 2021/02, but it fell back in the interval of 2020/04-2020/06, 2020/08-2020/09 and 2021/01-2021/02. The standard deviation of new deaths peaked at 260.16 in 2021/01.   
  
In our project, we will concern how monthly new cases and monthly new deaths change with the time in different regions.
  
```{r,message=FALSE, warning=FALSE,echo=F}
data_region<-data_region %>%  filter(month != "2021-03")
p1 <- data_region %>% 
  highlight_key(~WHO_region) %>%
  ggplot(aes(y = RNew_cases, x = month, color = WHO_region,
             )) + labs(title = "Time Series plot of new cases in different WHO regions")+geom_line(aes(group = WHO_region)) +ylab('WHO_region')+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))


ggplotly(p1, tooltip = c('x', 'group', 'text')) %>% 
  highlight(on = "plotly_hover", off = "plotly_doubleclick") %>% 
  layout(
    updatemenus = list(list(
      type = "buttons", 
      direction = "left", 
      xanchor = "left",
      yanchor = "top", 
      showactive = FALSE,
      buttons = list(list(
        method = "restyle", args = list("visible", "all"), label = "show all"),
        list(method = "restyle", args = list("visible", "legendonly"), label = "hide all")
      )
    ))
  )

```
  
The time series plot of new cases in different WHO regions shows the number of new cases of six WHO regions from 2020/01 to 2021/02. New cases in AFRO had two waves of slow growth and decreased after the growth. On the whole, the change of new cases in AFRO is smooth. New cases in WPRO changed more slowly than new cases in AFRO. New cases in EMRO also had two waves of slow growth and decreased after the growth. The increase of new cases in EMRO was more severe than in AFRO. The number of new cases in SEARO increased from 2020/01 to 2020/09 and decreased from 2020/09 to 2021/02. The number of new cases in EURO increased from 2020/01 to 2020/04 and from 2020/06 to 2020/11. The number of new cases in EURO decreased from 2020/04 to 2020/06 and from 2020/11 to 2021/02. The number of new cases in AMRO increased from 2020/01 to 2020/08 and from 2020/09 to 2021/01. The number of new cases in AMRO decreased from 2020/08 to 2020/09 and from 2021/01 to 2021/02.

```{r,message=FALSE, warning=FALSE,echo=F}
p2 <- data_region %>% 
  highlight_key(~WHO_region) %>%
  ggplot(aes(y = RNew_deaths, x = month, color = WHO_region,
             )) + labs(title = "Time Series plot of new deaths in different WHO regions")+geom_line(aes(group = WHO_region)) +ylab('WHO_region')+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))


ggplotly(p2, tooltip = c('x', 'group', 'text')) %>% 
  highlight(on = "plotly_hover", off = "plotly_doubleclick") %>% 
  layout(
    updatemenus = list(list(
      type = "buttons", 
      direction = "left", 
      xanchor = "left",
      yanchor = "top", 
      showactive = FALSE,
      buttons = list(list(
        method = "restyle", args = list("visible", "all"), label = "show all"),
        list(method = "restyle", args = list("visible", "legendonly"), label = "hide all")
      )
    ))
  )
```
  
The time series plot of new deaths in different WHO regions shows the number of new deaths of six WHO regions from 2020/01 to 2021/02. New deaths in AFRO had two waves of slow growth and decreased after the growth. New deaths in WPRO changed more slowly than new deaths in AFRO. New deaths in EMRO also had two waves of slow growth and decreased after the growth, while the increase of new deaths in EMRO was more severe than in AFRO. The number of new cases in SEARO increased from 2020/01 to 2020/09 and decreased from 2020/09 to 2021/02, which is consistent with the increase of new cases in SEARO. The number of new deaths in EURO increased from 2020/01 to 2020/04 and from 2020/08 to 2021/01. The number of new deaths in EURO decreased from 2020/04 to 2020/08 and from 2021/01 to 2021/02. The number of new deaths in AMRO increased from 2020/01 to 2020/08 and from 2020/11 to 2021/01. The number of new deaths in AMRO decreased from 2020/08 to 2020/11 and from 2021/01 to 2021/02.
  
```{r,message=FALSE, warning=FALSE,echo=F}

world <- map_data("world");
worldplot <- ggplot() +
  geom_polygon(data = world, aes(x=long, y = lat, group = group)) 

c_data$Country[c_data$Country =="United States"]<-"USA"
c_data$Country[c_data$Country =="S. Sudan"]<-"South Sudan"
c_data$Country[c_data$Country =="The United Kingdom"]<-"UK"
c_data$Country[c_data$Country =="Republic of Korea"]<-"South Korea"
c_data$Country[c_data$Country =="Cabo Verde"]<-"Cape Verde"
c_data$Country[c_data$Country =="	Czechia"]<-"Czech Republic"
c_data$death_rate[is.nan(c_data$death_rate)]<-0
covid.today<- c_data %>% 
   mutate(region=Country)


covid.today.world<- inner_join(world, covid.today, by = "region")
#cumulative cases
fig.map1  <- ggplot() +
  geom_polygon(data = covid.today.world, aes(x=long, y = lat, group = group,fill=Cumulative_cases)) + 
  labs(title ="The number of cumulative cases over countries")+
  coord_fixed(1.3)
plotly::ggplotly(fig.map1)

```
In the plot of cumulative cases over different countries from 2020/01/03 to 2021/2/28, the number of cumulative cases in the United States was the highest, which is 28174978. India had the second highest cumulative cases 11096731 and Brazil had the third highest cumulative cases 10455630. Then we explore the geometric distribution of the cumulative deaths. 

```{r,message=FALSE, warning=FALSE,echo=F}
# cumulative deaths
fig.map2  <- ggplot() +
  geom_polygon(data = covid.today.world, aes(x=long, y = lat, group = group,fill=Cumulative_deaths)) + 
  labs(title ="The number of cumulative deaths over countries")+
  coord_fixed(1.3)
plotly::ggplotly(fig.map2)

```
In the plot of cumulative deaths over different countries from 2020/01/03 to 2021/2/28, the number of cumulative deaths in the United States was the highest. Brazil had the second highest cumulative deaths and India had the third highest cumulative deaths. The United States, India, and Brazil are all have a large population, so it is unreasonable to judge the severity of Covid-19 by the cumulative cases and cumulative deaths. To eliminate the effect of population base, we use death rate to see which country was in the worst situation.

```{r,message=FALSE, warning=FALSE,echo=FALSE}
# death rate
fig.map3  <- ggplot() +
  geom_polygon(data = covid.today.world, aes(x=long, y = lat, group = group,fill=death_rate)) + 
  labs(title ="The death rate over countries")+
  coord_fixed(1.3)
plotly::ggplotly(fig.map3)
```
In the plot of death rate over different countries from 2020/01/03 to 2021/2/28, Yemen had the highest death rate 0.2780. Mexico had the second highest death rate 0.0888, and China had the third highest death rate 0.0475.
  
```{r, message=FALSE, warning=FALSE,echo= F}
data %>% 
  group_by(month) %>% 
  e_charts(Country, timeline = TRUE) %>%
  e_map(New_cases) %>% 
  e_visual_map(min = 0, max = max(data$New_cases), type = 'piecewise',
               splitList = list(
                 list(min = 10000000),
                 list(min = 1000000, max = 10000000),
                 list(min = 100000, max = 1000000),
                 list(min = 10000, max = 100000),
                 list(min = 1000, max = 10000),
                 list(min = 100, max = 1000),
                 list(min = 0, max = 100)
               )) %>% 
  e_tooltip(trigger = "item", formatter = e_tooltip_choro_formatter()) %>% e_title(text = 
  'The number of new cases over time',
          left = 'center',
          top = '0.8%',
          textStyle = list(fontWeight = 'bold', fontSize = 20))
```
Between 2020/01 and 2020/01, the covid was most heavy in China. The number of new cases grew rapidly from 9724 to 69669 for a month. So China compulsively requires people to cooperate in Covid-19 prevention. And the effect is very significant, the number of new cases dropped to 3152 on 2020/03. However, the covid-19 spread rapidly around the world’s cha as Spain had 108364 number of new cases in March. From 2020/03 to 2021/01, the map became redder and redder, especially North America and Asia & Pacific except China. Most countries had a decrease in February 2021.   
  
```{r, message=FALSE, warning=FALSE,echo=FALSE}
data %>% 
  group_by(month) %>% 
  e_charts(Country, timeline = TRUE) %>%
  e_map(New_deaths) %>% 
  e_visual_map(min = 0, max = max(data$New_deaths), type = 'piecewise',
               splitList = list(
                 list(min = 100000),
                 list(min = 10000, max = 100000),
                 list(min = 1000, max = 10000),
                 list(min = 100, max = 1000),
                 list(min = 0, max = 100)
               )) %>% 
  e_tooltip(trigger = "item", formatter = e_tooltip_choro_formatter()) %>% e_title(text = 
  'The number of new deaths over time',
          left = 'center',
          top = '0.8%',
          textStyle = list(fontWeight = 'bold', fontSize = 20))
```
Between 2020/01 and 2020/02, the number of new deaths grew rapidly from 213 to 2625 for a month. After China compulsively required people to cooperate in Covid-19 prevention，the number of new cases drop to 476 in 2020/03. At that time, in the United States, Spain, France and some countries,  more than 2000 people died because of Covid-19. The covid-19 spread rapidly around the regions, Canada and Mexico turned red as the United States, South America were still safe compared with other regions in April. From 2020/05 to 2021/02, the number of death cases is changing in different countries. However, Russian, Brazil, United States and so on were always red, the number of people who died in these countries more than ten thousand each month.

```{r, message=FALSE, warning=FALSE,echo=FALSE}
#pie chart of RNew_cases over time
 p1<-data_region %>% 
  group_by(month) %>% 
  e_charts(WHO_region, timeline = TRUE, playInterval =1000) %>%
  e_pie(RNew_cases) %>% 
  e_labels(fontSize = 13, formatter = '{b}: {d}%') %>% 
  e_legend(type = 'scroll', selector = TRUE,
           selectorPosition = 'start',
           selectorLabel = list(fontSize = 15, fontWeight = 'bolder')) %>%
  e_title(text = 
  'New cases', left = 'center',
          top = '5%',
          textStyle = list(fontWeight = 'bold', fontSize = 20))%>%e_tooltip()
p1
```
  
Between 2020/01 and 2020/01, the covid was most heavy in Regional Office for Western Pacific (WPRO), more than 97% of new cases were in WPRO. On 2020/03, the number of new cases decreased rapidly in WPRO. However, the Regional Office for Europe (EURO) accounts for 65.1% of all regions with 459642 number of new cases. The covid-19 spread rapidly around all regional offices. Until 2021/02, Regional Office for Americas (AMRO) account for most parts of all regions, besides 2020/10, Europe (EURO) had 8081456 new cases, accounting for 47.1%.
  
```{r, message=FALSE, warning=FALSE,echo=FALSE}
#pie chart of RNew_deaths over time
p2<-data_region %>% 
  group_by(month) %>% 
  e_charts(WHO_region, timeline = TRUE, playInterval =1000) %>% e_pie(RNew_deaths) %>% 
  e_labels(fontSize = 13, formatter = '{b}: {d}%') %>% 
  e_legend(type = 'scroll', selector = TRUE,
           selectorPosition = 'start',
           selectorLabel = list(fontSize = 15, fontWeight = 'bolder')) %>% e_title(text = 
  'New deaths', left = 'center',
          top = '5%',
          textStyle = list(fontWeight = 'bold', fontSize = 20))%>% e_tooltip()
p2

#p3<-p2 %>% e_connect('map') 
#e_arrange(p1, p3, cols=2)
```
  
Between 2020/01 and 2020/01, the covid was most heavy in Regional Office for Western Pacific (WPRO), more than 97% of new deaths were in WPRO. In 2020/03, the number of new deaths decreased rapidly in WPRO from 2648 to 811. Then, Regional Office for Europe (EURO) accounts for 80.91% of all regions with 29435 number of new cases On 2020/03 and account 58.37% of all regions with 108540 number of new cases On 2020/04. Later, The covid-19 spread rapidly around the Regional Office for the Americas (AMRO). Until 2021/02, Regional Office for Americas (AMRO) account for the most of all regions, besides 2020/10 and 2020/11, Regional Office for Europe (EURO) had more new deaths than Regional Office for Americas (AMRO).

Overall, the trend of new cases and new deaths seemed similarly.





# Inferential analysis 

## ANOVA model

As we discuss before, we can pay attention on the relationship between monthly new cases and monthly new deaths,but first, we will concern whether there is any differences existing in new cases and new deaths in each month between different regions. We can use two-way ANOVA model to answer this question.    

```{r,echo=FALSE,warning=FALSE}
plotmeans(RNew_cases~WHO_region,data = data_region)
plotmeans(RNew_deaths~WHO_region,data = data_region)
```
  
We can see that both monthly new cases and monthly new deaths show differences in different regions and according to the previous analyze, we know that the new cases and new deaths also show differences in different month, so we can try to conduct ANOVA.   
  
$Y_{itk}=\mu_{..}+\alpha_i+\tau_t++\epsilon_{itk}$     
$Y_{it}$ stands for the new cases in $i$ region in $t$ month. $\alpha_i$ stands for the effect from different regions and $\tau_t$ stands for effect from different months.
```{r,echo=FALSE}
aov1 = aov(RNew_cases~WHO_region+month,data = data_region)
```
$Z_{itk}=\nu_{..}+\alpha_i+\tau_t++\epsilon_{itk}$     
$Z_{it}$ stands for the new deaths in $i$ region in $t$ month. $\alpha_i$ stands for the effect from different regions and $\tau_t$ stands for effect from different months.
```{r,echo=FALSE}
aov2 = aov(RNew_deaths~WHO_region+month,data = data_region)
```
  
And we can conduct the test to show whether the coefficients of these two model are significant:    
$H_{\alpha0}:\alpha_i = 0, \ for \ all \ i$  
$H_{\alpha1}:not \ all \ \alpha_i \ are \ 0$  
$H_{\tau0}: \tau_t = 0, \ for \ all \ t$  
$H_{\tau1}:not \ all \ \tau_t \ are \ 0$  
We can check the summary of the first ANOVA model:
```{r,echo=FALSE}
summary(aov1)
```
and the summary of the second ANOVA model:
```{r,echo=FALSE}
summary(aov2)
```


According to the summary of these two ANOVA models,both the p-values of F test of month and WHO_region are very small, so we can consider that both of WHO_region and moth are statistically significant in these two models, which means that the number of new cases and new deaths are different in different month and different regions.    

  
## Regression model
As we discussed in ANOVA part, the number of monthly new cases and monthly new deaths are different in different regions and according to the time series plots, we can find that the new cases and new deaths show the similar trends in each region, so we may consider to analyze the relationship between monthly new cases and monthly new deaths treating regions as dummy variable. First, we can check the scatter plot between monthly new cases and monthly new deaths.   
  
```{r,echo=FALSE}
ggplot(data = data_region)+
  geom_point(mapping = aes(x= RNew_cases,y = RNew_deaths,color = WHO_region ))
```
In this plot, we can see that for different region, the rate of change between of new cases and new deaths seem to be different, but if we draw the regression line for each region, both of them seem to pass the origin, so it may imply us to regress the model without interception and  treat the regions as dummy variables which will affect the slope of the model.    
  
The model will be:    
$Z_j = \beta_1 Y_jR_1 +\beta_2 Y_jR_2+\beta_4 Y_jR_3+\beta_4 Y_jR_4+\beta_5 Y_jR_5+\beta_6 Y_jR_6+\epsilon_j$, $\epsilon_j \sim N(0,\sigma^2),i.i.d$    
$Z_j$ stands for the monthly new deaths, $Y_j$ stands for the monthly new cases and $R_1,R_2,...R_{6}$ are the dummy variables representing different WHO regions.    
We can check the summary of this model:    
```{r,echo=FALSE}
fit1 = lm(RNew_deaths~RNew_cases+RNew_cases:factor(WHO_region),data = data_region)
summary(fit1)
fit_ad = lm(RNew_deaths~RNew_cases:factor(WHO_region),data = data_region)
summary(fit_ad)
fit3 = fit_ad = lm(RNew_deaths~RNew_cases:factor(WHO_region)-1,data = data_region)
summary(fit3)
```
For all coefficients of this model, we can conduct the test:    
$H_0: \beta_1= \beta_2=...\beta_{6} = 0$   
$H_1: not \ all \ \beta_i \  are \ 0$  
  
According to the summary of the model, the p-value of F-statistic is very small, which means that we can reject $H_0$ and consider that this model is statistically significant on the whole.And for single coefficient, we can see that most of the coefficients are significant too.
  
In our model, we do not consider the influence of time. Now we can use mixed effect regression treating month as random effect. And we can check the summary of this mixed effort model.
```{r,echo=FALSE,warning=FALSE}
fit2 = lmer(RNew_deaths~RNew_cases:factor(WHO_region)+(1|month),data = data_region)
summary(fit2)
```
We can get the similar conclusion. 


# Sensitivity analysis 
We can check the basic assumption of our model and do some diagnostics.
```{r,echo=FALSE}
par(mfrow = c(1,2))
plot(fit_ad,which = c(1,2))
```


```{r,echo=FALSE}
par(mfrow = c(1,2))
plot(fit3, which = c(1,2))
```
  
According to the plots, this model can not fit the basic assumptions really well,especially the normality,so we may consider do some transformation on our model.To be convenient,we can remove the data which new deaths are 0.And according to the boxcox, we can use log(RNew_deaths). We can see the summary and the plots of the transformed model.
Final model : $logZ_j = \beta_1 Y_jR_1 +\beta_2 Y_jR_2+\beta_4 Y_jR_3+\beta_4 Y_jR_4+\beta_5 Y_jR_5+\beta_6 Y_jR_6+\epsilon_j, \ Z_j >0$
```{r,echo=FALSE,message=FALSE}
library(MASS)
data_change = data_region %>% filter(RNew_deaths != 0)
fit_change = lm(RNew_deaths~RNew_cases:factor(WHO_region)-1,data = data_change)
boxcox(fit_change)
data_change$RNew_deaths = log(data_change$RNew_deaths)
fit_final = lm(RNew_deaths~RNew_cases:factor(WHO_region)-1,data = data_change)
summary(fit_final)
par(mfrow = c(1,2))
plot(fit_final,which = c(1,2))
```
  
We can see that all of the coefficients are significant and according to the QQ plot,the model seems to be normality, but it shows a line in Residuals vs Fitted, which may imply that there exist heteroscedasticity or other linear relationship in our model. 







#  Causal interpretation 



# Discussion 



# Acknowledgement {-}



# Reference {-}



# Session info {-}



```{r}
sessionInfo()
```